---
title: "p8105_hw5_zl3119"
author: "Zheyan"
date: "11/11/2021"
output: github_document
---

```{r, include = FALSE}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "95%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 2

Data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

### Create and tidy data

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time; Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary


```{r Read data, warning=FALSE, message=FALSE}
data_path = 'data'
file_list = list.files(data_path)


extract_ID_arm = function(filename) {
    filefront = strsplit(filename, split = '.', fixed = TRUE)[[1]][1]
    # extract info about arm and ID
    arm = strsplit(filefront, split = '_', fixed = TRUE)[[1]][1]
    ID = as.integer(strsplit(filefront, split = '_', fixed = TRUE)[[1]][2])
    # read file
    df_temp = read_delim(paste(data_path, filename, sep = '/'), delim = ',')
    df_temp %>%  
         mutate(arm = arm,
         ID = ID) 
}


df = map_df(list.files(data_path), extract_ID_arm) %>% 
  arrange(arm, ID) %>% 
  relocate(arm, ID) %>% 
  pivot_longer(week_1:week_8, names_to = 'week', names_prefix = 'week_',values_to = 'value') %>% 
  mutate(week = as.factor(week))

# show data
head(df,16) %>% 
  knitr::kable()

```

### Visualization

Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r}
df %>% 
  ggplot(aes(x = week, y = value, group = paste0(ID, arm), color = arm)) +
  geom_line()
```


Two groups are similar at the first week. Paticipants in the experiment group have increasing value while those from control group have fluctuate values. Paticipants from  experiment grouphave higher value at last two weeks.

# Problem 3

```{r generate missing data, include = FALSE}
set.seed(10)

iris_with_missing = iris %>% 
  # every column has 20 missing values
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```


For numeric variables, you should fill in missing values with the mean of non-missing values
For character variables, you should fill in missing values with "virginica"

```{r}

# define the impute values
impute = 
  map_df(iris_with_missing, ~mean(.x, na.rm = TRUE)) %>% 
  mutate(Species = "virginica")
  
# copy an iris set to impute on
iris_imputed = iris_with_missing

# impute by columns 
for (col in colnames(iris_with_missing)){
  colvalues = iris_imputed[col]
  iris_imputed[col] = 
    colvalues %>% 
    replace(is.na(colvalues), impute[[col]])
}

print(paste0('Missing value counts:',sum(is.na(iris_imputed))))

# show imputed dataframe
head(iris_imputed,12) %>% 
  knitr::kable()

```






